Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-12-16 PAPER/Paola_2/ms_ver3.tex
Modified text 1

To obtain the SiO emission maps and corresponding
spectra, two additional free parameters are required along with other
inputs described in section~\ref{sec:radtrans} . They are
the fractional abundance profile of SiO and the angle of inclination with
respect to line of sight denoted by $\phi$. Section~\ref{ssec:sioabun}
describes all the profiles used for the present study. $\phi$ = 90$^{\circ}$ indicates that
the outflow is in the plane of the sky. 
Additionally, we have used two other angles of inclination apart from the
plane of the sky, i.e, $\phi = 45^{\circ}, 60^{\circ}$ to compare
results with observations. The runs with different fractional
abundance profiles are described in table~\ref{tab:result2}. A
parameter introduced here is $\delta$ which is essentially the ratio
of shock velocity v$_{\rm s}$ and physical jet velocity v$_{\rm
  jet}$ such that its value depends on density contrast $\eta$
(see Eq.~\ref{eq:shockjetvel}).

\begin{table*}
\centering
\caption{Summary from parameter runs. The parameter $\delta$, is the ratio of
  shock to jet velocity and it depends on the density contrast $\eta$
  as shown in eq.~\ref{eq:shockjetvel}}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile
  $\delta = 1$} & \multicolumn{2}{|c|}{Gaussian Profile $\delta =
  1$}\\
\hline\hline
&&& $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi10 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow10 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm10 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm2 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab10 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab2 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol10 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol3 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

\begin{table*}
\centering
\caption{Summary of radiative transfer runs with different SiO fractional abundance
  profiles for dynamical simulation with molecular cooling and
  $\eta=3$. The integrated intensity assuming a single disk
  observation with a beam width of $15\,\arcsec$ is listed along with the
  corresponding spectral width.}
\begin{tabular}{l | l | l | l | l}
\hline
Profile & $\delta$ = v$_{\rm s}$/v$_{\rm jet}$ & $\int T_{\rm MB}dV$
[K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$]\\ 
\hline\hline
Top Hat & 1.0 & 3.3 & 14.0 \\
Top Hat & 1/$(1.0 + \eta(z)^{-0.5})$ & 4.9 & 14.0 \\
Gaussian & 1.0 & 0.66
& 12.0 \\
Gaussian & 1/$(1.0 + \eta(z)^{-0.5})$ & 1.1
& 12.0 \\
%Functional & 100.0 & $(1.0 + \eta(z)^{-0.5})$ &
%& \\
%Functional & 100.0 & $(1.0 + \eta(z)^{-0.5})$ &
%& \\
\hline
\end{tabular}
\label{tab:result2}
\end{table*}

@@

The cooling in the reference jet is via molecular cooling prescription, whereby
the initial hydrogen fractions are evolved using appropriate rate equations and with cooling 
contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr), i.e, density, temperature and velocity obtained from this 
run are given as inputs to the radiative transfer code to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in the plane of the sky ($\phi =
90^{\circ}$). A gaussian profile without any upper cutoff is used for the fractional abundance
of SiO. Further, the shock velocity which determines
the fractional abundance, v$_{\rm s}$, is less than the jet velocity in the
flow, such that their ratio $\delta$ is given by
equation~\ref{eq:shockjetvel}, where $\eta$ is the density contrast as a
function of height from the base of the jet.

@@

The dynamical states for each of this cooling prescriptions are then 
given as inputs to the radiative transfer code to obtain integrated J = 2-$>$1 SiO emission
maps shown in figure~\ref{fig:coolcmpsio}. The differences in dynamical
features are also reflected in the corresponding SiO emission
maps. These maps are obtained assuming an angle of inclination of
90$^\circ$ using the top hat abundance
profile with $\delta = 1$. They are then placed at a distance of
100$\,$pc and convolved with a beam with FWHM of 2.0\arcsec in order
to create the maps shown. 
The emission from the jet with atomic
cooling is coming from the conical jet head and younger knots closer
to base of the flow. Additionally, features in the shell at Z $\sim$
60, 75 also show evidence of faint emission. In case of the jet with
power-law cooling the emission mostly arises from the
internal knots. 
The emission coming from jets with
tabulated and molecular cooling resembles that observed in young molecular outflows.
Also in these outflows, strong emission arises from the
condensation formed at the bow shock due to thermal
instability. 
%

@@

For the reference run, the effect of the shock in the dynamical evolution of
the hydrogen fraction is presented in figure~\ref{fig:Hmolform}. The top
panel of the figure shows the hydrogen number density in the jet. 
The fraction of all hydrogen forms are shown in bottom
panels of figure~\ref{fig:Hmolform} along two representative regions in
the jet marked with the arrows. The left bottom panel shows the
vertical distribution of temperature, molecular and atomic fraction at the
interface of the knot and the ambient medium. The interaction of the
knot with the medium raises that temperature to 1200\,K and
accumulates the matter so that the density increases up to
10$^{6}$ cm$^{-3}$. Behind the shock as the material cools, atoms
combine together to form molecules as seen in the increase of fH2 from
0.44 to 0.52. Further away from the
shock, atomic and molecular hydrogen tend to reach a quasi equilibrium as their
fractions reach a value of 0.5. The ionized fraction is
extremely low in this region due to low temperatures. However at the
bow-shock, temperatures rise up to 20000\,K. Molecular
hydrogen species are destroyed, while ionized hydrogen shows an
increase in its fraction 
(figure~\ref{fig:Hmolform}). The peak in ionized fraction of 0.15
coincides with the peak in temperature profile as expected. The
molecular fraction shows a considerable dip from 0.3 to 0.03 at 
this temperature before rising sharply in the ambient molecular
medium. 
%

@@

The spectra at three different positions in the flow are shown in the
bottom panel of figure~\ref{empvspec90} (they are marked with green
crosses in top left panel). The knot closer to the base is the brightest
showing a peak intensity of 25\,K. The
intensity decreases and reaches to about 10\,K close to the bow shock.
The line widths seen are typically around 5-10 km s$^{-1}$.
These line widths increase substantially as the angle of inclination
decreases. Figure \ref{empvspec45} show the
emission map, PV diagram and spectra for the same reference run but
with angle of inclination of 45$^{\circ}$
respectively. The line 
widths now are typically of the order of 20 km s$^{-1}$. 
The PV diagrams in these runs are not longer symmetric. Instead, they show a distinct {\it zig-zag}
pattern at each of the knots and region close to the bow shock.

@@

\subsection{SiO Abundance profile}
\label{ssec:sioabunem}
The fractional abundance of SiO has been assumed to vary with shock velocity in
a simplistic manner (see sec~\ref{ssec:sioabun}). In
figure~\ref{fig:diffabunem}, we compare the maps for J = 2-$>$1
emission at an inclination angle of 90$^{\circ}$ for three different abundance profiles and two different ratios of shock to
jet velocities, $\delta$. There is a striking difference with regards to
emission from the internal knots in these images. Using the top hat profile
and accounting for shock speeds (i.e. $\delta < 1$) results in
emission from  all the internal knots in the J=2-$>$1 line. However, some of
these knots are not observed when using the same top-hat profile but
assuming the shock velocity to be same as the jet velocity (i.e.,
$\delta$ = 1). Similar qualitative characteristics are seen in case of a gaussian
abundance profile. In particular, the case with $\delta$ = 1 only
produces emission from the knot closest to the bow shock, while the internal
knots do not show any appreciable emission. The step function and
gaussian step (not shown) with $\delta = 1$ produces very similar
features as that of the top hat profile with $\delta < 1$.
Further, the emission from
the knot closer to the bow shock varies considerably with different
profiles and value of $\delta$. It is the brightest for the case with
a top-hat profile and $\delta < 1$. The peak emission and line widths at this knot
are listed in table~\ref{tab:result2}.
%

@@

The dependence of emission on SiO abundance profile is expected due to
the distribution of jet velocity obtained from dynamical
simulations. We see that the velocity of internal knots lie around
70-90\,km\,s$^{-1}$, while the pulsed jet was injected with a mean of
100\,km\,s$^{-1}$. The knots slow down during the evolutionary
phase as they interact with the ambient medium. Interestingly, younger
knots closer to the base of the jet are brighter compared to older
ones further away from the jet (see panels 1 and 3 of
fig~\ref{fig:diffabunem}). This is attributed to the fact that the
ambient medium has a density gradient that goes as $\sim
z^{-2}$. Thus, the younger knots suffer the most deceleration closer
to the base of flow. This fact is taken into account when the shock
velocity is consistently calculated using the density contrast and
using a value of $\delta <$ 1. This process is further validated by
the lack of emission from internal knots in panel 3 of the
fig~\ref{fig:diffabunem}. 
%

Additionally, the internal knots show their distinct signatures in
form of {\em Hubble wedges} as seen in the PV diagrams for these
different profiles (see
~\ref{ssec:emspecpv}). Figure~\ref{fig:pvdiffabun} shows a zoomed in
version of the bottom four internal knots in form of a PV diagram for these
different abundance profiles. As expected, the signatures of knots 
is missing for the case with gaussian profile with $\delta$ =
1. Further, the wedges formed in panels 1, 2 and 5 of the figure in
general are slightly more extended as compared to those
formed in panel 4. This is indeed because of the
discontinuous nature of the top-hat abundance profile as the SiO is
enhanced to a maximum abundance for all velocities between 20 and 100
km\,s$^{-1}$, which is not the case in a more continuous gaussian
distribution.

@@

\section{ALMA view}
\label{sec:ALMAview}
%
In order to see the interior structure of the molecular outflow and
test the model predictions, high
resolution interferometric observations are needed. 
We have performed synthetic Atacama Large Millimeter
Array (ALMA) observations using the Common Astronomy Software
Applications package (CASA). The output from the radiative transfer
were scaled to a distance of 300pc and used as the sky models for observations of
the J=2-1, J=5-4 and J=8-7 lines at frequencies of 86.85, 217.10 and
347.33$\,$GHz, which fall in ALMA bands 3, 6 and 7, respectively. The observations
were simulated using cycle 1 ALMA for 30 minutes total integration
time, in the configurations c32-6, c32-4 and c32-3 for the three
lines, giving beam sizes of 0.72'', 0.47'' and 0.58'' and
sensitivities of 0.05, 0.07 and 0.09$\,$mJy/beam for the three lines
(as calculated by the ALMA online sensitivity calculator) with
velocity resolutions of $\sim$1$\,$km/s.
 Figure~\ref{fig:almafig}
presents the ALMA prediction obtained from our reference run. The
left panel shows the integrated emission of SiO J = 2-1 line while the right
shows the PV diagram for the same line and over laid are
the contours of SiO J = 5-4 and J = 8-7. The contour levels are marked
on the color bar of each panel.

@@

In IRAS 04166+2706, \cite{SantiagoGarcia:2009p13972} have shown 
that the EHV CO(2-1) emission is mapping a jet-like
feature that consists of a collection of
discrete peaks symmetrically placed on both outflow lobes. This symmetry
indicate that such EHV peaks might arise from events that took place
near the central source and have since propagated in the flow
\citep{Bachiller:1990p11196, Tafalla:2011p14051}. The dynamical 
model in conjunction with proper radiative transfer calculations presented here 
agrees very well with the above scenario. The {\em{bullets}} in our
work are injected into the flow in a sinusoidal manner along with a
collimated atomic jet. These pulsating ejections do interact with the medium
via shocks and exciting high velocity gas. Including the cooling associated with
molecules and the H$_{2}$ chemistry allows us to
consistently identify the regions where molecular hydrogen is formed
and disassociated (see fig.~\ref{fig:Hmolform}). The SiO emission that we
obtain from our radiative transfer calculations is associated with
regions slightly behind where the shocked H$_{2}$ gas is present
as observed in case of L1157 molecular outflow
\cite{Gueth:1998p14058}. A contour map of SiO
emission obtained from the reference run is shown in
figure~\ref{fig:ltconts}. Further, we find SiO emission coming from velocities 
of the order of 40-60 km\,s$^{-1}$ (see fig
~\ref{empvspec45}). Additionally, we see a distinct saw-tooth pattern
in the PV diagrams from our models (see figs. \ref{empvspec45} and
\ref{fig:almafig}). These predictions from our model very well
resembles the characteristic features of EHV SiO emission 
seen in the majority of the outflows \citep{SantiagoGarcia:2009p13972,
  Tafalla:2011p14051, Arce:2013p14902}.
%

@@

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$, assuming optically thin emission. 
They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines. Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see
fig.~\ref{grid}). These are the regions where the bulk of the SiO
emission is present, with fractional abundances between 10$^{-6}$ to 10$^{-8}$.
The line profiles shapes obtained from our study fit reasonable well with that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. Interestingly, in our models, the line profiles are equally broadened for low as well
as high excitation lines. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination
$\phi$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec to simulate observed integrated
intensities of L1448 bullets observed with large beams from JCMT and
IRAM. Intensities predicted from our model (see table~\ref{tab:result2}) lie within a factor of two
from the observed values (Table 2 of \citealt{Nisini:2007p13128}).
%

The variation of integrated single dish emission with upper transition
levels J$_{\rm up}$ for different abundance profiles is shown in the
right panel of fig.~\ref{fig:lineratio}. It is clear that the emission
obtained using a top hat profile is brighter than the one obtained
with a Gaussian SiO profile. However, curves obtained from
both these profiles show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. These curves helps to to predict the
hydrogen number density and temperature in outflows. Physical
conditions in outflows can also be studied by estimation of line ratios.
Values $\approx$ 1 for line ratios SiO(8-7)/(5-4) and SiO(5-4)/(2-1) have been
found in HH212 \citep{Cabrit:2007p13804,
  Lee:2008p13697} and also for molecular outflows from massive
young stellar object IRAS 17233-3606 (\citealt{Leurini:2013p13165}).
The local velocity gradient (LVG) slab modeling with optically thin approximation were unable to produce
these line ratios. 
In the present non-LTE radiative transfer modeling, no assumption of optical
depths or photon escape probability are prescribed and line ratios are
self consistently calculated and are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. 
%Further, we have estimated the optical
%depth in regions emitting SiO and find them to be optically thick with
%an optical depth $\tau \sim 10$. Thus, assuming optically thin SiO emission
%from molecular outflows especially from shocked dense knots would
%only provide upper limits. 
In summary, the shape, ratio and peak intensity of the spectra
obtained from our model fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in the gas phase via shocks.

@@

\subsection{Slow Component and Grain Chemistry}
%
We have presented a model that is able to explain the spectral characteristics
of the SiO EHV component seen
in several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in presence of
appropriate cooling. The steady state temperature, density and velocities obtained from dynamical
simulation is used as input in the non-LTE radiative transfer code to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in gas phase from grain-grain and grain-mantle
collisions exists, their focus is mainly on weaker shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800, VanLoo:2013p16573}. 
The pulsating jet propagation model presented here have shock speeds
reaching up to 100 km\,s$^{-1}$. We have used few
simplified SiO abundance profiles as a function of shock
velocity. Though the shape of the profile is uncertain, the upper and
lower limits used are backed by observational evidences. Ideally, for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper 
as this could imply to include dust grains and sputtering 
as well as grain-grain collisions (i.e. an MHD multi-fluid approach).  
%

>>>>

To obtain the SiO emission maps and corresponding
spectra, two additional free parameters are required along with other
inputs described in section~\ref{sec:radtrans} . They are
the fractional abundance profile of SiO and the angle of inclination with
respect to line of sight denoted by $\phi$. Section~\ref{ssec:sioabun}
describes all the profiles used for the present study. $\phi$ = 90$^{\circ}$ indicates that
the outflow is in the plane of the sky. 
Additionally, we have used two other angles of inclination apart from the
plane of the sky, i.e, $\phi = 45^{\circ}, 60^{\circ}$ to compare
results with observations. The runs with different fractional
abundance profiles are described in section~\ref{ssec:sioabunem}

\begin{table*}
\centering 
\caption{Summary of parameter runs for different cooling prescriptions and density contrasts using two abundance profiles.}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile} & \multicolumn{2}{|c|}{Gaussian Profile}\\
\hline\hline
&&& $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi10 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow10 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm10 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm2 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab10 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab2 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol10 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol3 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

%\begin{table*}
%\centering
%\caption{Summary of radiative transfer runs with different SiO fractional abundance
%  profiles for dynamical simulation with molecular cooling and
%  $\eta=3$. The integrated intensity assuming a single disk
%  observation with a beam width of $15\,\arcsec$ is listed along with the
%  corresponding spectral width.}
%\begin{tabular}{l | l | l | l | l}
%\hline
%Profile & $\delta$ = v$_{\rm s}$/v$_{\rm jet}$ & $\int T_{\rm MB}dV$
%[K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$]\\ 
%\hline\hline
%Top Hat & 1.0 & 3.3 & 14.0 \\
%Top Hat & 1/$(1.0 + \eta(z)^{-0.5})$ & 4.9 & 14.0 \\
%Gaussian & 1.0 & 0.66
%& 12.0 \\
%Gaussian & 1/$(1.0 + \eta(z)^{-0.5})$ & 1.1
%& 12.0 \\
%%Functional & 100.0 & $(1.0 + \eta(z)^{-0.5})$ &
%%& \\
%%Functional & 100.0 & $(1.0 + \eta(z)^{-0.5})$ &
%%& \\
%\hline
%\end{tabular}
%\label{tab:result2}
%\end{table*}

@@

The cooling in the reference jet is via molecular cooling prescription, whereby
the initial hydrogen fractions are evolved using appropriate rate equations and with cooling 
contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr), i.e, density, temperature and velocity obtained from this 
run are given as inputs to the radiative transfer code to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in the plane of the sky ($\phi =
90^{\circ}$). A Gaussian profile with a step is used for the fractional abundance
of SiO.

@@

The dynamical states for each of this cooling prescriptions are then 
given as inputs to the radiative transfer code to obtain integrated J = 2-$>$1 SiO emission
maps shown in figure~\ref{fig:coolcmpsio}. The differences in dynamical
features are also reflected in the corresponding SiO emission
maps. These maps are obtained assuming an angle of inclination of
90$^\circ$ using the top hat abundance
profile. They are then placed at a distance of
16.7$\,$pc and convolved with a beam with FWHM of 2.0\arcsec in order
to create the maps shown. 
The emission from the jet with atomic
cooling is coming from the conical jet head and younger knots closer
to base of the flow. Additionally, features in the shell at Z $\sim$
60, 75 also show evidence of faint emission. In case of the jet with
power-law cooling the emission mostly arises from the
internal knots. 
The emission coming from jets with
tabulated and molecular cooling resembles that observed in young molecular outflows.
Also in these outflows, strong emission arises from the
condensation formed at the bow shock due to thermal
instability. 
%

@@

For the reference run, the effect of the shock in the dynamical evolution of
the hydrogen fraction is presented in figure~\ref{fig:Hmolform}. The top
panel of the figure shows the hydrogen number density in the jet. 
The fraction of all hydrogen forms are shown in bottom
panels of figure~\ref{fig:Hmolform} along two representative regions in
the jet marked with the arrows. The left bottom panel shows the
vertical distribution of temperature, molecular and atomic fraction at the
interface of the knot and the ambient medium. The interaction of the
knot with the medium raises that temperature to 1000\,K and
accumulates the matter so that the density increases up to
10$^{6}$ cm$^{-3}$. Behind the shock as the material cools, atoms
combine together to form molecules as seen in the increase of fH2 from
0.44 to 0.52. Further away from the
shock, atomic and molecular hydrogen tend to reach a quasi equilibrium as their
fractions reach a value of 0.5. The ionized fraction is
extremely low in this region due to low temperatures. However at the
bow-shock, temperatures rise up to 20000\,K. Molecular
hydrogen species are destroyed, while ionized hydrogen shows an
increase in its fraction 
(figure~\ref{fig:Hmolform}). The peak in ionized fraction of 0.15
coincides with the peak in temperature profile as expected. The
molecular fraction shows a considerable dip from 0.3 to 0.03 at 
this temperature before rising sharply in the ambient molecular
medium. 
%

@@

The spectra at three different positions in the flow are shown in the
bottom panel of figure~\ref{empvspec90} (they are marked with green
crosses in top left panel). The knot closer to the base and the region near the bow shock 
shows a peak integrated intensity of 15-20\,K. Whereas, the integrated 
intensity decreases and reaches to about 5\,K in the middle of the jet.
The line widths seen are typically around 5-10 km s$^{-1}$.
These line widths increase substantially as the angle of inclination
decreases. Figure \ref{empvspec45} show the
emission map, PV diagram and spectra for the same reference run but
with angle of inclination of 45$^{\circ}$
respectively. The line 
widths now are typically of the order of 20 km s$^{-1}$. 
The PV diagrams in these runs are not longer symmetric. Instead, they show a distinct {\it zig-zag}
pattern at each of the knots and region close to the bow shock.

@@

\subsection{SiO Abundance profile}
\label{ssec:sioabunem}
The fractional abundance of SiO has been assumed to vary with jet flow velocity in
a simplistic manner (see sec~\ref{ssec:sioabun}). In
figure~\ref{fig:diffabunem}, we compare the maps for J = 2-$>$1
emission at an inclination angle of 90$^{\circ}$ for four different abundance profiles. 
There is a striking difference with regards to
emission from the internal knots in these images. Using the step profile results in
emission from  all the internal knots in the J=2-$>$1 line. However, some of
these knots are not observed when using the same top-hat profile. 
Similar qualitative characteristics are seen in case of a Gaussian
abundance profile. In particular, the case without a step beyond 60\,km\,s$^{-1}$
produces emission from the region closest to the bow shock and the lower most knot, 
while the internal knots do not show any appreciable emission. While, the 
Gaussian profile with a step produces very similar
features as that of the step profile effectively making the initial rising part 
of the curve redundant. 
%

@@

The dependence of emission on SiO abundance profile is expected due to
the distribution of jet velocity obtained from dynamical
simulations. We see that the velocity of internal knots lie around
80-90\,km\,s$^{-1}$, as the pulsed jet was injected with a mean of
100\,km\,s$^{-1}$. The knots slow down during the evolutionary
phase as they interact with the ambient medium. Interestingly, younger
knots closer to the base of the jet are brighter compared to older
ones further away from the jet (see panels 1 and 3 of
fig~\ref{fig:diffabunem}). This is attributed to the fact that the
ambient medium has a density gradient that goes as $\sim
z^{-2}$. Thus, the younger knots suffer the most deceleration closer
to the base of flow. This process is further validated by
the decreasing abundance profile giving rise to 
lack of emission from internal knots in panel 3 of the fig~\ref{fig:diffabunem}. 
%

Additionally, the internal knots show their distinct signatures in
form of {\em Hubble wedges} as seen in the PV diagrams for these
different profiles (see
~\ref{ssec:emspecpv}). Figure~\ref{fig:pvdiffabun} shows a zoomed in
version of the bottom four internal knots in form of a PV diagram for these
different abundance profiles. As expected, the signatures of knots 
is missing for the case with Gaussian profile without a step.
Further, the wedges formed in panels 1, 2 and 4 of the figure in
general are slightly more extended as compared to those
formed in panel 3. This is indeed because of the
discontinuous nature of the top-hat abundance profile as the SiO is
enhanced to a maximum abundance for all velocities between 20 and 100
km\,s$^{-1}$, which is not the case in a more continuous Gaussian
distribution.

@@

\section{ALMA view}
\label{sec:ALMAview}
%
In order to see the interior structure of the molecular outflow and
test the model predictions, high
resolution interferometric observations are needed. 
We have performed synthetic Atacama Large Millimeter
Array (ALMA) observations using the Common Astronomy Software
Applications package (CASA). The output from the radiative transfer
were scaled to a distance of 300\,pc and used as the sky models for observations of
the J=2-1, J=5-4 and J=8-7 lines at frequencies of 86.85, 217.10 and
347.33$\,$GHz, which fall in ALMA bands 3, 6 and 7, respectively. The observations
were simulated using cycle 2 ALMA for 25 minutes total integration
time, in the configurations c32-6, c32-4 and c32-3 for the three
lines, giving beam sizes of 0.72'', 0.47'' and 0.58'' and
sensitivities of 0.05, 0.07 and 0.09$\,$mJy/beam for the three lines
(as calculated by the ALMA online sensitivity calculator) with
velocity resolutions of $\sim$1$\,$km/s.
 Figure~\ref{fig:almafig}
presents the ALMA prediction obtained from our reference run. The
left panel shows the integrated emission of SiO J = 5-4 line while the right
shows the PV diagram for the same line and over laid are
the contours of SiO J = 2-1 and J = 8-7. The contour levels are marked
on the color bar of each panel.

@@

In IRAS 04166+2706, \cite{SantiagoGarcia:2009p13972} have shown 
that the EHV CO(2-1) emission is mapping a jet-like
feature that consists of a collection of
discrete peaks symmetrically placed on both outflow lobes. This symmetry
indicate that such EHV peaks might arise from events that took place
near the central source and have since propagated in the flow
\citep{Bachiller:1990p11196, Tafalla:2011p14051}. The dynamical 
model in conjunction with proper radiative transfer calculations presented here 
agrees very well with the above scenario. The {\em{bullets}} in our
work are injected into the flow in a sinusoidal manner along with a
collimated atomic jet. These pulsating ejections do interact with the medium
via shocks and exciting high velocity gas. Including the cooling associated with
molecules and the H$_{2}$ chemistry allows us to
consistently identify the regions where molecular hydrogen is formed
and disassociated (see fig.~\ref{fig:Hmolform}). The SiO emission that we
obtain from our radiative transfer calculations is associated with
regions slightly behind where the shocked H$_{2}$ gas is present
as observed in case of L1157 molecular outflow
\cite{Gueth:1998p14058}. A contour map of SiO
emission obtained from the reference run is shown in
figure~\ref{fig:ltconts}. Further, we find SiO emission coming from velocities 
of the order of 40-70 km\,s$^{-1}$ (see fig
~\ref{empvspec45}). Additionally, we see a distinct saw-tooth pattern
in the PV diagrams from our models (see figs. \ref{empvspec45} and
\ref{fig:almafig}). These predictions from our model very well
resembles the characteristic features of EHV SiO emission 
seen in the majority of the outflows \citep{SantiagoGarcia:2009p13972,
  Tafalla:2011p14051, Arce:2013p14902}.
%

@@

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$, assuming optically thin emission. 
They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines. Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see
fig.~\ref{grid}). These are the regions where the bulk of the SiO
emission is present, with fractional abundances between 10$^{-6}$ to 10$^{-8}$.
The line profiles shapes obtained from our study fit reasonable well with that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. Interestingly, in our models, the line profiles are equally broadened for low as well
as high excitation lines. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination
$\phi$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec to simulate observed integrated
intensities of L1448 bullets observed with large beams from JCMT and
IRAM. Intensities predicted from our reference model, $\int T_{\rm MB} dV$ = 5.76 for SiO(2-1), 
lie within a factor of two from the observed values (Table 2 of \citealt{Nisini:2007p13128}).
%

The variation of normalized integrated single dish emission with upper transition
levels J$_{\rm up}$ for the reference run is shown in the
right panel of fig.~\ref{fig:lineratio}. The curve obtained 
show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. Such a curve is used for kinematic studies and 
helps to predict the hydrogen number density and temperature in outflows. Physical
conditions in outflows can also be studied by estimation of line ratios.
Values $\approx$ 1 for line ratios SiO(8-7)/(5-4) and SiO(5-4)/(2-1) have been
found in HH212 \citep{Cabrit:2007p13804,
  Lee:2008p13697} and also for molecular outflows from massive
young stellar object IRAS 17233-3606 (\citealt{Leurini:2013p13165}).
The local velocity gradient (LVG) slab modeling with optically thin approximation were unable to produce
these line ratios. 
In the present non-LTE radiative transfer modeling, no assumption of optical
depths or photon escape probability are prescribed and line ratios are
self consistently calculated and are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. 
%Further, we have estimated the optical
%depth in regions emitting SiO and find them to be optically thick with
%an optical depth $\tau \sim 10$. Thus, assuming optically thin SiO emission
%from molecular outflows especially from shocked dense knots would
%only provide upper limits. 
In summary, the shape, ratio and peak intensity of the spectra
obtained from our model fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in the gas phase via shocks.

@@

\subsection{Slow Component and Grain Chemistry}
%
We have presented a model that is able to explain the spectral characteristics
of the SiO EHV component seen
in several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in presence of
appropriate cooling. The steady state temperature, density and velocities obtained from dynamical
simulation is used as input in the non-LTE radiative transfer code to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in gas phase from grain-grain and grain-mantle
collisions exists, their focus is mainly on weaker shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800, VanLoo:2013p16573}. 
The pulsating jet propagation model presented here have flow speeds
reaching up to 120 km\,s$^{-1}$. We have used few
simplified SiO abundance profiles as a function of jet
velocity. Though the shape of the profile is uncertain, the upper and
lower limits used are backed by observational evidences. Ideally, for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper 
as this could imply to include dust grains and sputtering 
as well as grain-grain collisions (i.e. an MHD multi-fluid approach).  
%
